---
title: "disasteRs"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    df_print: default
    number_sections: no
vignette: >
  %\VignetteIndexEntry{disasteRs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(disasteRs)
```

## Summary

Our package includes data from the International Disaster Database data acquired from the [EM-DAT](https://www.emdat.be/). The users can import the occurrence and impacts of over 26,000 mass disasters worldwide from 1900 to the present day for data analysis. \

The package includes three primary functions:

### 1. **`plot_disaster_map_world()`** and **`plot_disaster_map_sea()`**

- **Purpose:** Displays the frequency of natural disasters on a world or Southeast Asia map.
- **Output:** Choropleth maps where darker regions indicate higher frequencies of disasters.
- **Use Case:** Ideal for identifying which regions are most prone to disasters.

### 2. **`disaster_month()`**

- **Purpose:** Summarizes and visualizes the months with the most frequent disaster occurrences in a specified region or country.
- **Output:** Bar plots and summaries of the number of disasters starting in each month.
- **Use Case:** Helps analyze seasonal trends in disaster occurrences.

### 3. **`analyze_disasters()`**

- **Purpose:** Summarizes and visualizes the human impacts (e.g., deaths, injuries, affected populations) of a specific disaster type for a given year.
- **Output:** Bar graphs, tables, or both, breaking down disaster impacts.
- **Use Case:** Assesses the human cost of specific disasters.

## Data Sets Included

- `emdat_world`: the occurrence and impacts of over 16,000 mass disasters worldwide from 2000 to 2024

## Basic Usage

```{r, message=FALSE, warning=FALSE}
library(disasteRs)
```

### Data Wrangling

- Using the built-in dataset 'disaster_data'emdat_world' included in the package, the user can easily analyze the most frequent disaster types worldwide for a given year:

```{r example, message=FALSE, warning=FALSE}
library(dplyr)

most_frequent_disasters <- emdat_world |>
  filter(start_year == 2023) |>
  group_by(disaster_type) |>
  summarize(count = n()) |>
  arrange(desc(count))

head(most_frequent_disasters)
```
